---
title: "Palmetto Binary Logistic Regression"
description: |
  In this project, I explored Florida palmetto data through visualizations and binary logistic regression. Data source: Abrahamson, 2019.
author:
  - name: Michelle Shteyn
date: 02-04-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

# Overview

In this report, I explore Florida palmetto data collected by Abrahamson from 1981 to 2017 to examine the feasibility of using 4 explanatory variables (height, canopy length, canopy width, and green leaves) to classify whether a palmetto is one of two species: *Serenoa repens* or *Sabal etonia*. I used the following methods:  

- Conducting exploratory data visualizations to examine relationships among explanatory variables using `GGally::ggpairs()`  
- Performing binary logistic regression using plant height, canopy length, canopy width and green leaves as predictor variables to understand how they relate to probability of a plant being *Serenoa repens* or *Sabal etonia*  
- Producing a finalized table showing the number and percentage of *Serenoa repens* and *Sabal etonia* palmettos correctly or incorrectly classified by my binary logistic regression model including 4 predictor variables and using a 50% cutoff

### Initial Data Wrangling:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE, warning=FALSE}
#Attach required packages
library(here)
library(tidyverse)
library(broom)
library(GGally)
library(jtools)
library(kableExtra)

#Read in data
palm <- read_csv(here("data","palmetto.csv")) %>% 
  mutate(species = as.factor(species)) %>% #Change species to a factor variable
  mutate(species = case_when(
    species==1~"S. repens", #Code 1's as S. repens
    species==2~"S. etonia"  #Code 1's as S. etonia
  ))
```

### Explore Differences between Two Palmetto Species: 

```{r, warning=FALSE,message=FALSE}
###Exploration with `ggpairs`
palm %>% 
  select(height,length,width,green_lvs, species) %>% 
  ggpairs(aes(color = species))+
  labs(title = "Differences in Height, Canopy Length, Canopy Width, and Green Leaves\nfor S. etonia and S. repens Palmetto Species, 1981 - 2017" )+
  theme_bw()
```

**Figure 1**. Exploring relationships among explanatory variables height, length, width, and number of green leaves of two dominant palmetto species of south-central Florida, ongoing at 5-year intervals. Source: Abrahamson, 2019.

This visualization demonstrates that, for both Palmetto species, the variables of height, length, and width are strongly related; however, the distributions in number of green leaves for the two species are quite different, with *S. repens'* distribution being more platykurtic than *S. etonia's*.


```{r, warning=FALSE,message=FALSE}
ggplot(data = palm, aes(x = green_lvs)) +
  geom_histogram(aes(fill=species), show.legend = FALSE) +
  facet_wrap(~species) +
  theme_bw()+
  labs(title = "Distributions in Number of Green Leaves for S. etonia and S. repens\n Palmetto Species, 1981 - 2017",
       x = "Number of Green Leaves",
       y = "Counts of Palmettos")
```

**Figure 2**. Frequency distributions describing the number of Palmetto trees which fall in different categories for number of green leaves, for both *Serenoa repens* and *Sabal etonia* species. Source: Abrahamson, 2019. 

Focusing on this distinction between the two species identified in the more comprehensive visualization above, this graph shows that, for *S. etonia* species, most palmettos contain around 3 leaves, while for *S. repens* species, most palmettos contain around 8 leaves.


### Binary Logistic Regression:

**Table 1**. Coefficients for the predictor variables (plant height, canopy length, canopy width and green leaves) that relate to the *log odds* of the *S. etonia* outcome:

```{r, message=FALSE, warning=FALSE}
#Conduct blr
palm1 <- read_csv(here("data","palmetto.csv")) %>% #Convert `species` to factor
  mutate(species = as.factor(species)) 

palm1_blr <- glm(species ~ height + length + width + green_lvs, 
                            data = palm1, 
                            family = "binomial")

# Get a tidy version w/ broom: 
blr_tidy <- broom::tidy(palm1_blr)


blr_tidy %>% 
  kable(col.names = c("Explanatory Variable",
                      "Coefficient",
                      "Std. Error for Coefficient",
                      "T-Statistic",
                      "P Value"),
        digits = c(3, 3, 3))%>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")
```


### Find actual probabilities associated with a palmetto being S. repens or S. etonia

**Table 2**. Number and percentage of *S. repens* and *S. etonia* palmettos correctly or incorrectly classified by our binary logistic regression model using 4 predictor variables (plant height, canopy length, canopy width and green leaves), using a 50% cutoff. Source: Abrahamson, 2019.

```{r, message=FALSE,warning=FALSE}
blr_fitted <- palm1_blr %>% 
    broom::augment(type.predict = "response") %>% 
  mutate(species = as.numeric(species)) %>% 
  select(-.rownames) %>% 
  round(digits = 3)

blr_fitted <- blr_fitted %>% 
  mutate(species = case_when(
    species==1~"S. repens",
    species==2~"S. etonia"
  )) 

blr_fitted_1<-blr_fitted %>% 
  select(species, .fitted) %>% 
  mutate(final_classification = case_when(
    .fitted >= .50 ~ "S. etonia",              #Since S. repens is the reference variable, a probability > 50% classifies it as S. etonia
   .fitted < .50 ~ "S. repens")) %>% 
  mutate(correct = case_when(
    species=="S. repens" & final_classification=="S. repens" ~ "Correct",
    species=="S. repens" & final_classification=="S. etonia" ~ "Incorrect",
    species=="S. etonia" & final_classification=="S. etonia" ~ "Correct",
    species=="S. etonia" & final_classification=="S. repens" ~ "Incorrect",
  ))

blr_fitted_1 %>% 
  group_by(species, correct) %>% 
summarize(count = n()) %>% 
  pivot_wider(names_from = correct, values_from = count) %>% 
  mutate(percent_correct = Correct/(Correct+Incorrect)*100) %>% 
    kable(col.names = c("Plant Species",
                      "Number Correctly Classified",
                      "Number Incorrectly Classified",
                      "Percent Correctly Classified")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")
```
# Key Takeaways
The binary logistic regression model using 4 predictor variables (plant height, canopy length, canopy width and green leaves) was successful over 90% of the time in correctly classifying whether a palmetto species is *S. etonia* or *S. repens*.

## Data Source
Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

**More information and metadata:**
https://portal.edirepository.org/nis/metadataviewer?packageid=edi.317.1


