---
name: Michelle
surname: Handy
position: "Ph.D. Candidate"
address: "Department of Psychological & Brain Sciences, University of California, Santa Barbara"
phone: 973 641 6454
www: michelleshteynhandy.com
email: "michelleshteyn@gmail.com"
github: michelleshteyn1
linkedin: michelleshteynhandy
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::moderncv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where,
  "B.S. with Honors", "2016", "University of Maryland, College Park", "Psychology",
  "M.A.", "2020", "University of California, Santa Barbara", "Psychological & Brain Sciences",
  "Ph.D.", "2021", "University of California, Santa Barbara", "Psychological & Brain Sciences"
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```
# Theses

```{r}
tribble(
  ~Category, ~Year, ~Title, ~Department,
  "Master's Thesis:","2020","Regional Extreme Weather Concern and its Relation to Support for Environmental Action","Department of Psychological & Brain Sciences (advisor: David Sherman, Ph.D.)",
  "Honor's Thesis:","2016","Examining the Effects of Combined Construal on Task Persistence","Department of 			Psychology, University of Maryland (mentor: Richard Yi, Ph.D.)"
) %>% 
  detailed_entries(Category, Year, Title, Department)
```


# Professional Memberships

```{r}
tribble(
  ~Org, ~Year,
  "Society for Personality and Social Psychology (SPSP)","2016-present",
  "Association for Psychological Science (APS)","2016-present",
  "American Psychological Association (APA)","2016-present",
  "Phi Beta Kappa Society","2016-present",
) %>% 
 detailed_entries(Org, Year)
```

# Research Experience

## Paid Positions

```{r}
tribble(~Position, ~Year, ~Project, ~Department,
        "Graduate Student Researcher ($1,000)","2020","as part of “Health Psychology & COVID-19 Interview Project”","APS Teaching Fund Microgrants program (PI: Dr. David Sherman)",
        "Graduate Student Researcher ($6,185)","2020","as part of LVLWGB, NSF Grant","(PI: Dr. Lisa Leombruni with Co-PI: Dr. Heather Hodges)",
        "Graduate Student Researcher ($7,137)","2019","as part of “Sociocultural Determinants of Human Motives”","NSF Grant (PI: Dr. David Sherman with Co-PI: Dr. Heejung Kim)",
        "Graduate Student Researcher ($5,223)","2019","as part of “Gift Fund; Debris Flow, Project Code KE538”","(PI: Dr. Ed 		Keller)",
        "Graduate Student Researcher ($2000)","2018","as part of “Promoting HPV Vaccination among College Students Via Health Message Framing”","UC Health Psychology Consortium Seed Grant, 2017, $9,994 (PI: Dr. David 			Sherman with Co-PI: Dr. Kate Sweeny)"
        ) %>% 
  detailed_entries(Position,Year,Project,Department)
```

# Publications

1. Sherman, D. K., **Shteyn, M. F.**, Han, H., & Van Boven, L. (2020). The exchange between citizens and elected officials: A social psychological approach to lobbying for climate policy. *Behavioural Public Policy*.
2. Sherman, D.K., Updegraff, J., **Handy, M.S.**, Eom, K., & Kim, H. (in press). Beliefs and social norms as precursors of environmental support: The joint influence of collectivism and socioeconomic status. Personality & Social Psychology Bulletin. 

# Manuscripts under Editorial Review or In Preparation

1. Anderson, S.E., **Handy, M.S.**, Gehlbach, H., & Virlar-Knight, D. (in preparation). Psychological distance and the effect of connecting extreme weather events to climate change.  
2. Binning, K.R., Conrique, B.I., Areguin, M., Brick, C., Ehret, P., **Handy, M.S.**, Updegraff, J.A., & Sherman, D. K. (in preparation). Health messaging in a diverse society: Effects of gain and loss frames on health behaviors 	among latina/os in the U.S.  
3. **Handy, M.S.**, Block, D.M., Kim, H., Knowles, E.D., Mildenberger, M., & Sherman, D.K. (in preparation). Regional 	extreme weather concern and its relation to support for environmental action. 

# Published Conference Abstracts or Presentations

1. **Shteyn, M.F.** (2020, February). *Regional Extreme Weather Concern and its Relation to Support for Environmental Action*. Single Presenter talk at SPSP’s Symposium on “Harnessing Social Psychology to Address Social Issues and Public 	Policy,” New Orleans, LA.  
2. **Shteyn, M.F.** (2019, August). *Psychological Proximity to Extreme Weather and its Relation to Regional Climate Change Beliefs and Policy Support*. Data blitz talk presented at SPF 2019: Big Data in Personality and Social Psychology, St. Louis, MO.  
3. **Shteyn, M.F.** (2019, February). *Immediacy and Psychological Distance: Effects on Environmental Action*. Poster presentation at the SPSP Sustainability Psychology Preconference, Portland, OR.  
4. **Shteyn, M.F.** (2017, January). *Discussing the Null Effects of Construal Level on Task Persistence*. Poster presentation at the SPSP Self-Regulation Preconference, San Antonio, TX.

# Research Awards and Academic Honors

```{r}
tribble(~Category, ~Year, ~Title, ~Department,
        "Graduate Travel Award Winner","2020","2020 SPSP Convention","",
        "Registration Award","2019","SPSP Summer Psychology Forum: Big Data","",
        "Senior Marshal","2016","2016 Psychology Department Commencement Ceremony","University of Maryland",
        "Summer Scholars Award","2015","2015 BSOS Dean's Research Initiative ($2000)","University of Maryland",
        "College Park Scholars Citation","2014","Life Sciences Program","",
        "Outstanding Citizenship Award","2014","College Park Scholars","",
        "President's Scholarship ($20,000)","2012-2016","University of Maryland",""
  
) %>% 
  detailed_entries(Category, Year, Title, Department)
```

# Teaching

## Teaching Assistant

```{r}
tribble(
  ~Class, ~Year,
  "Health Psychology","2018, 2021",
  "Instroduction to Psychology","2018, 2020",
) %>% 
 detailed_entries(Class, Year)
```

## Lab Instructor

```{r}
tribble(
  ~Class, ~Year,
  "Statistical Methods in Psychological & Brain Sciences","2020",
  "Addressing Social Problems with Psychological Science","2019",
  "Laboratory in Research Methods","2018",
  "Laboratory in Advanced Research Methods","2018",
  "Laboratory in Advanced Research Methods in Social Psychology","2017,2020"
 
) %>% 
 detailed_entries(Class, Year)
```

## Mentoring

```{r}
tribble(~Position, ~Year, ~Department,~Description,
       "Mentor of 3 research assistants and leader of weekly academic reading group","Present","","",
       "McNair Scholar Summer Mentor","2018","UCSB Academic Research Consortium","Mentored a first-generation, underrepresented student in her applications to Ph.D. programs",
       "Appointed Career Panel Speaker","2016","Freshman Colloquim","University of Maryland College Park Scholars Life Sciences",
       "Peer Mentor for Freshman Class","2013-2014","College Park Scholars Life Sciences",""
       
) %>% 
  detailed_entries(Position, Year, Department, Description)
  
```

## Guest Lectures

```{r}
tribble(~Title, ~Year, ~Class, ~Department,
        "Immediacy and Psychological Distance: Effects on Environmental Action","2019","Guest lecture	for Addressing Social Problems with Psychological Science (PSY160)","Psychological & Brain Sciences 	Department, UC Santa Barbara. *(also cited above under Published Conference Abstracts and Presentations)",
        "Health Psychology and the Environment","2019","Guest lecture for Health Psychology (PSY101)","UC Santa Barbara") %>% 
  detailed_entries(Title,Year, Class, Department)
```

# Professional Activities

```{r}
tribble(~Position,~Year,
        "Reviewer for 2 papers for the Journal of Behavioural Public Policy","2019") %>% 
  detailed_entries(Position, Year)
```

# Invited Talks

1. Leombruni, L., Hodges, H., **Shteyn, M.F.**, & Brush, A. (2020, October). E*nhancing Experiential Digital Learning: Exploring the Impact of Interactive & Narrative-Driven Media on Informal STEM Learning in Kids & Young Adults*. 	Presentation to PBS NOVA Polar Extremes.  
2. **Shteyn, M.F.** (2020, February). *Regional Extreme Weather Concern and its Relation to Support for Environmental 	Action*. Social Psychology Area Meeting presentation, Psychological & Brain Sciences Department, UC Santa 	Barbara.*(also cited above under Published Conference Abstracts and Presentations)  
3. **Shteyn, M.F.** (2020, January). *Construal Level and Work Motivation*. Lab meeting presentation, Management and Organization Department, USC Marshall School of Business.  
4. **Shteyn, M.F.** (2019, May). *Immediacy and Psychological Distance: Effects on Environmental Action*. Graduate student mini-convention presentation, Psychological & Brain Sciences Department, UC Santa Barbara.  
5. **Shteyn, M.F.** (2016, April). *Examining the Effects of Combined Construal on Task Persistence*. Poster presentation at the Annual Department of Psychology Poster Presentations, University of Maryland, College Park, MD. *(also cited above under Published Conference Abstracts and Presentations)

# Service

```{r}
tribble(~Position, ~Year, ~Org, ~Location,
        "Chair of Symposium","2020","SPSP, “Harnessing Social Psychology to Address Social Issues and Public Policy”","New Orleans, LA") %>% 
  detailed_entries(Position, Year, Org,Location)
```
































