---
title: "Multiple Linear Regression: Black-Throated Blue Warblers"
description: |
  In this project, I explored data on black-throated blue warblers through exploratory visualizations and multiple linear regression. Data source: Rodenhouse et al., 2019.
author:
  - name: Michelle S. Handy
    url: {}
date: 02-19-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: 'hide'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#Attach packages
library(tidyverse)
library(here)
library(stargazer)
library(naniar)
library(broom)
```

# Overview

This report analyses data collected by Dr. Rodenhouse and colleagues at Hubbard Brook Experimental Forest (1986 - 2010). Below, we explore relationships between selected black-throated blue warbler characteristics and mass using the following methods:

- Creating exploratory ggplots 
- Using multiple linear regression to describe relationships between bird mass and explanatory variables
- Justifying variable/model selection based on diagnostic plots, overall model fit, AIC, and conceptual considerations
- Presenting outcomes of which variables predict bird mass in a finalized regression table


```{r}
#Read in data, replace the value -99.99 with NA

birds <- read.csv(here("data","bt_warblers.csv")) %>% 
  replace_with_na(replace = list(bill_depth = -99.99)) %>% 
  replace_with_na(replace = list(bill_culmen = -99.99)) %>% 
  replace_with_na(replace = list(mass= -99.99)) %>% 
  replace_with_na(replace = list(tarsus1 = -99.99)) %>% 
  replace_with_na(replace = list(wing_r = -99.99)) 

```

# Visually explore the data

### 1: Wing length vs. bird mass

- Point color indicating sex
- Faceted by bird age at capture

```{r, message=FALSE,warning=FALSE}
ggplot(data = birds, aes(x=wing_r, y = mass)) + 
  geom_point(aes(color = sex)) + 
  facet_wrap(~age_banded) + 
  labs(x = "Wing Length (mm)",
       y = "Body Mass (g)",
       title = "Black-Throated Blue Warbler Bird Mass, as a Function of Wing Length")
```

### 2: Tarsus length vs. bird mass

- Point color indicating sex
- Faceted by elevation

```{r, message=FALSE,warning=FALSE}
ggplot(data = birds, aes(x=tarsus1, y = mass)) + 
  geom_point(aes(color = sex)) + 
  facet_wrap(~elevation) + 
  labs(x = "Tarsus Length (mm)",
       y = "Body Mass (g)",
       title = "Black-Throated Blue Warbler Bird Mass, as a Function of Tarsus Length")
```

### 3: Bill depth vs. bird mass

- Point color indicating sex
- Faceted by bird age at capture

```{r, message=FALSE,warning=FALSE}
ggplot(data = birds, aes(x=bill_depth, y = mass)) + 
  geom_point(aes(color = sex)) + 
  facet_wrap(~age_banded) + 
  labs(x = "Bill Depth (mm)",
       y = "Body Mass (g)",
       title = "Black-Throated Blue Warbler Bird Mass, as a Function of Bill Depth")
```



```{r, include=FALSE}
#Use MLR describe relationships between bird mass & explanatory variables. Try different combinations of exploratory variables, then consider diagnostics, model fit, and AIC, to decide which model version to present. 

#Build a few possible models
bird_lm <- lm(mass ~ sex + age_banded + bill_depth + bill_culmen + tarsus1 + wing_r, data = birds)
bird_lm1 <- lm(mass ~ sex + bill_depth + bill_culmen + tarsus1 + wing_r, data = birds)
bird_lm2 <- lm(mass ~ sex + bill_depth + tarsus1 + wing_r, data = birds)
bird_lm3 <- lm(mass ~ sex + wing_r, data = birds)

#Examine overall model fit: they all seem around the same (R^2 ~.30)
summary(bird_lm)
summary(bird_lm1)
summary(bird_lm2)
summary(bird_lm3)

#Look at model diagnostics
plot(bird_lm) #One data point (332) seems to have large influence on this model (Cook's distance)
plot(bird_lm1)
plot(bird_lm2)
plot(bird_lm3) # 2 outliers (obs 332 & 247) here, concerning

#Compare the model AIC values
AIC(bird_lm, k = 2) #1197.157
AIC(bird_lm1, k = 2) #1193.2 - this is the lowest AIC
AIC(bird_lm2, k = 2) #1199.775
AIC(bird_lm3, k = 2) #2366.062

#Based on these 4 models only, the Akaike Information Criterion indicates that 'bird_lm1' is the best balance between fit & complexity. It also makes sense conceptually, as sex and different body measurements are all variables which tend to influence an animal's size.
```


# Finalized Regression Table

```{r,message=FALSE, warning=FALSE,echo=FALSE, results='asis'}
stargazer(bird_lm1, type = "html", 
  covariate.labels = c("Sex: Male","Bill Depth","Bill Culmen","Tarsus Length","Wing Length"),
  dep.var.labels = "Bird mass")
```
### Model Selection Process

4 possible explanatory models of bird mass were compared. These were: 

(1) Bird Mass ~ Sex + Age Banded + Bill Depth + Bill Culmen + Tarsus Length + Wing Length
(2) Bird Mass ~ Sex + Bill Depth + Bill Culmen + Tarsus Length + Wing Length
(3) Bird Mass ~ Sex + Bill Depth + Tarsus Length + Wing Length
(4) Bird Mass ~ Sex + Wing Length

- Overall model fit for the four models were quite similar, with the variables explaining around 30% of the variance in body mass. 
 - The Akaike Information Criterion indicates that **Model 2** is the best balance between fit & complexity. It also makes sense conceptually, as sex and various body measurements are all variables which tend to influence an animal's size.
 
 
# Interpreting the model results
```{r, include=FALSE}
bird_lm1_tidy <- tidy(bird_lm1)

# Return it:
bird_lm1_tidy
```
- **Sex:** If birds are consistent across all other variables, we expect a *male* bird to weigh `r round(bird_lm1_tidy$estimate[2],2)` g **less** than a female bird, on average.  
- **Bill Depth (mm):** The coefficient for bill depth (`r round(bird_lm1_tidy$estimate[3],2)`) is the average change we expect to see in body mass (grams) for a 1 mm increase in bill depth. 
- **Bill Culmen (mm):** The bill culmen coefficient (`r round(bird_lm1_tidy$estimate[4],2)`) is the average change we expect to see in body mass (grams) for a 1 mm increase in bill culmen. 
- **Tarsus Length (mm):** The tarsus length coefficient (`r round(bird_lm1_tidy$estimate[5],2)`) is the average change we expect to see in body mass (grams) for a 1 mm increase in tarsus length. 
- **Right Wing Length (mm):** The right wing length coefficient (`r round(bird_lm1_tidy$estimate[6],2)`) is the average change we expect to see in body mass (grams) for a 1 mm increase in right wing length. 


# Data Citation:

Rodenhouse, N., R. Holmes, S. Sillett, and M. Webster. 2019. Black-throated Blue Warbler capture histories, Hubbard Brook Experimental Forest, 1986 - 2010 ver 4. Environmental Data Initiative. [https://doi.org/10.6073/pasta/ea985022acfdb76cd7beeaf8da6c39b1]



















