---
title: "Rasters and Spatial Data Visualization: Aquatic Species"
description: |
  In this project, I explored raster data on cetacean species off the coast of California through spatial visualization. Data source: Kaschner et al., 2016.
author:
  - name: Michelle S. Handy
    url: {}
date: 02-19-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster) 
library(tidyverse)
library(here)
library(sf)
library(fasterize)
library(rnaturalearth)
library(sp)
library(janitor)
```

# Overview

This report explores the probability of occurrence of 35 cetacean species that can be found in the California Bight, based on relative environmental suitability. The raster data is subsetted from AquaMaps (Kaschner et al., 2016; [www.aquamaps.org](www.aquamaps.org)), where the extent of the rasters is 125째 W to 115째W and and 32째N to 38째N. I use the following methods below:

- Use `raster::stack()` to create a raster of species richness of cetacean species off the coast of California. The threshold above which a species is considered "present" is set at .6.
- Create a map of species richness on the California coast using `coord_sf()` to define the boundaries of the map to match the raster extent. 

# Loading and inspecting the data

Load in the CA cetaceans file. Read in multiple layers at once as a multi-layer `RasterStack` object.

``` {r, warning=FALSE,message = FALSE, results = 'hide', cache = TRUE}
my_files <- list.files(path = "/Users/michelleshteyn/Documents/Academic Quarters/ESM244/Week 7/a3_task2/ca_cetaceans", full.names = TRUE, pattern = ".tif") #GeoTIFF format is most common raster file type

cetacean_stack <- raster::stack(my_files)

#plot(cetacean_stack)


```

# Create a species richness raster of cetacean species

First, I set the probability threshold to .6 using `calc()` to reclassify each probability layer to a "presence" layer of either 0 (not present) or 1 (present). Then, I apply another `calc()` to find the sum of the species that exist in each cell.

```{r, warning=FALSE, message=FALSE, results = 'hide'}
is_present <- function(x, thresh = .6){ #Created a function to set the threshold
  y <- ifelse(x >= thresh, 1, NA)
  return(y)
}

species_richness <- calc(cetacean_stack, fun = is_present) #Applied the threshold function to the raster stack `cetacean_stack` to create a species richness raster

#Apply another calc() to determine how many species in each cell
species_richness1 <- calc(species_richness, fun = sum, na.rm = TRUE) #Need to make sure the threshold function was giving either a 0 or a 1 rather then an NA or 1. If I sum an object with an NA, the results will be NA. Typically should set na.rm = TRUE.

```


# Create a map using `ggplot()`

Need to convert the raster to a dataframe so that it is a suitable object for `ggplot()`. First, use the `raster::rasterToPoints()` function to convert the raster to a matrix of points. Then, use `as.data.frame()` to convert that matrix to a `data.frame` object.

```{r, cache = TRUE, results = 'hide', warning = FALSE, message = FALSE, fig.cap= "This map depicts the probability of occurrence, based on relative environmental suitability, of 35 cetacean species that can be found in the California Bight. Each unit of the map depicts the number of species that occurs in that particular location. The threshold value above which a species is considered present was set at 0.6. This raster data was subsetted from AquaMaps (Kaschner et al., 2016)."}
# Convert species_richness1 into a dataframe using `raster::rasterToPoints`
species_richness1_df <- raster::rasterToPoints(species_richness1) %>%
  as.data.frame() 

#Include a shapefile of the coastline for reference from the `rnaturalearth` package
california110<-ne_download(scale = 110, type = 'states', category = 'cultural',returnclass="sf")

ggplot() +
geom_raster(data = species_richness1_df, aes(x=x,y=y, fill=layer)) +
  geom_sf(data = california110) +
  coord_sf(expand = 0, xlim = c(-125,-115), ylim = c(32,38)) +
  scale_fill_gradient(low = 'white', high = 'slateblue4') +
  theme_minimal() +
  theme(panel.background = element_rect(fill = 'slateblue4'))+
  labs(x = "Latitude", y = "Longitude", title = "Map of Cetacean Species Richness off the California Coast", fill = "Number of Cetacean Species")

```

# Data citation

Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). *AquaMaps: Predicted range maps for aquatic species*. (www.aquamaps.org)


