---
title: "Agglomerative Hierarchical Clustering: Water Chemistry"
description: |
  In this project, I explored water chemistry data in the Santa Barbara Coastal drainage area through hierarchical clustering. Data source: Santa Barbara Coastal LTER and J. Melack. 2019.
author:
  - name: Michelle S. Handy
    url: {}
date: 03-01-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Overview

This report explores stream chemistry in the Santa Barbara Coastal drainage area using data (ongoing since 2000) provided by Santa Barbara Coastal LTER and Dr. John Melack. Below, I use hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site. To perform hierarchical clustering by site, 

- I will first make a data frame that has a single summary row per site (based on means from all observations at that site).
- Then, I will calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering.


```{r, message=FALSE, warning=FALSE}
#Attach required packages
library(tidyverse)
library(here)

# Packages for cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

# Read in data and do some wrangling

```{r, message=FALSE, warning=FALSE}
stream <- read_csv(here("data","sbc_lter_registered_stream_chemistry.csv")) 

#Convert -999 values to NA
stream <- stream %>% 
na_if(-999)

#Make a data frame that has a single summary row per site. 
stream_summary <- stream %>% 
  group_by(site_code) %>% #Group by site
  summarise(mean_nh4_uM = mean(nh4_uM, na.rm = TRUE), #Calculate mean observation at that site
           mean_no3_uM = mean(no3_uM, na.rm = TRUE),
           mean_po4_uM = mean(po4_uM, na.rm = TRUE),
           mean_tdn_uM = mean(tdn_uM, na.rm = TRUE),
           mean_tdp_uM = mean(tdp_uM, na.rm = TRUE),
           mean_tpc_uM = mean(tpc_uM, na.rm = TRUE),
           mean_tpn_uM = mean(tpn_uM, na.rm = TRUE),
           mean_tpp_uM = mean(tpp_uM, na.rm = TRUE),
           mean_tss_mgperLiter = mean(tss_mgperLiter, na.rm = TRUE),
           mean_spec_cond_uSpercm = mean(spec_cond_uSpercm, na.rm = TRUE)) 

#Note: a few of these in the summary table says "NaN", which means "not a number." NaN implies a result that cannot be calculated for whatever reason, or is not a floating point number. In this data frame, it's because certain sites had all NAs for those chemistry measurements, and it is not possible to take a mean of a bunch of NAs. 
```

# Scale the data

```{r, message=FALSE, warning=FALSE}
# Scale the numeric variables (columns 2:12)
stream_scaled <- stream_summary %>% 
  select(2:11) %>% 
  scale()
  
# Update to add rownames (site code) from stream_summary
rownames(stream_scaled) <- stream_summary$site_code

# The rownames are now the site code (this is useful for visualizing)
```


# Calculate Euclidean distance

Use the `stats::dist()` function to find the Euclidean distance in multivariate space between the different observations (sites).

```{r, message=FALSE, warning=FALSE}
# Compute dissimilarity values (Euclidean distances):
euc_distance <- dist(stream_scaled, method = "euclidean")

```


# Perform hierarchical clustering by complete linkage with `stats::hclust()`

Next, I use the `stats::hclust()` function to perform hierarchical clustering on the matrix of euclidean distances (`euc_distance`) using complete linkage. 

In complete linkage, clusters are merged by the smallest maximum distance between two observations in distinct clusters.

```{r, message=FALSE, warning=FALSE}
# Hierarchical clustering (complete linkage)
hc_complete <- hclust(euc_distance, method = "complete" )

# Plot it (base plot):
# plot(hc_complete, cex = 0.6, hang = -1)

```


# Plot dendrogram using `ggplot()`

Using `ggdendrogram()`, a `ggplot` wrapper. 

```{r, message=FALSE, warning=FALSE}
ggdendrogram(hc_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(x = "Stream Site",
       y = "",
       title = "Multivariate Clustering for Water Chemistry by Site, Santa Barbara, CA")

```

# Data and metadata source

**SBC LTER:** Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000  
**Creators:** Santa Barbara Coastal LTER, & Melack, John M  
**Citation:** Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. (https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174.) 
